Supporting Information: Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes.

### 2020-06-10 Ben_Manuscripts/stochastic_transport/response_to_reviewers.tex

  % BJC: easy to specify force field.
  % BJC: what to say about polarizable force fields? Focus on methods? Maybe throw in some 
  % references to solute transport with polarizable force fields?
  % finite size effects -- ?? sufficiently long pores in z?
  % BJC: no gaurantee polarizable would be good since there aren't any out of the box 
  % polarizable force fields

  \item \textit{\textbf{Figure 5 page 25:} \\
		First, the symbols within the figure are very small and barely readable, and probably not
		according to authors guidelines for figures. Second, the caption, as with the other captions of
		this manuscript, contain a lot of interpretation that normally are placed in the main text. This
		might be a particular style of the authors that has its benefits, however in this specific case the
		caption fails at explaining what exactly this figure is showing. The way I interpret this figure is:
		the symbol on the y-axis marks a general parameter and the boldface symbol marks the
		parameter shown in the histogram. Am I correct? If so, I would strongly recommend modifying
		y-axis and the caption to make this clear.}

		We thank the reviewer for their constructive feedback. The fontsize in the figure is indeed too small.
		We have modified the plots in the figure to have larger, more readable font and we took this opportunity
		to increase the font size on Figures 6--8 as well.

		The reviewer is correct in their interpretation of the figures. For clarity, we modified the y axes as
		well as the caption in order to further aid the reader. We also made efforts to cut down on the interpretation
		that is already presented in the main text. We made similar modifications to Figure 6 since it is an
		analogous figure. Our caption changes are presented below.

		This comment caused us to reconsider the detail in the captions of other figures as well. We feel
		the level of the detail is justified in most cases. However, we decided to remove some redundant
		information from Figure 13, as shown below.

		Figure 5, original caption:

		\begin{quote}

		  The parameters of the one mode model reveal differences in dynamics
	      between solutes. (a) We parameterized Gaussian,
	      $\mathcal{N}(\sigma)$, and L\'evy stable, $L(\sigma, \alpha_h)$,
	      distributions to describe solute hop lengths. We assume the mean
	      ($\mu$) to be zero for these distributions and no
	      skewness ($\beta = 0$) in the L\'evy stable distributions. High
	      values of $\sigma$ and lower values of $\alpha_h$ result in larger
	      hops. (b) We parameterized a pure power law, $P(\alpha)$, and a
	      truncated power law, $P_T(\alpha, \lambda)$, distribution to describe
	      solute dwell times. Lower values of $\alpha$ lead to heavier power
	      law tails and higher values of $\lambda$ truncate the distribution at
	      lower dwell times. (c) Finally, we parameterized the hop
	      autocorrelation function, $\gamma(H)$, to describe the degree of
	      correlation between hops. Simulations with higher values of $H$ display
	      behavior closer to the Brownian limit.

		\end{quote}

		Figure 5, modified caption:

		\begin{quote}

		  The parameters of the one mode model reveal differences in dynamics
	      between solutes. (a) We parameterized Gaussian, $\mathcal{N}(\sigma)$, 
	      and L\'evy stable, $L(\sigma, \alpha_h)$, distributions to describe
	      solute hop lengths. 
		  % BJC: added	      
	      Each bar represents the value of a single parameter,
	      highlighted in bold, of the associated hop distribution.
		  % BJC: removed	      
	      % We assume the mean ($\mu$) to be zero for these distributions and no
%	      skewness ($\beta = 0$) in the L\'evy stable distributions. 
	      In general, higher values of $\sigma$ and lower values of $\alpha_h$ 
	      result in larger hops. 
	      (b) We parameterized 
	      % BJC: removed. I think this was left over from when the pure power law was still in the paper
	      % a pure power law, $P(\alpha)$, and 
	      a truncated power law distribution, $P_T(\alpha, \lambda)$,  to describe
	      solute dwell times. 
		  % BJC: not sure if should reiterate the bold highlights here.	      
	      Lower values of $\alpha$ lead to heavier power
	      law tails and higher values of $\lambda$ truncate the distribution at
	      lower dwell times. (c) Finally, we parameterized the hop
	      autocorrelation function, $\gamma(H)$, to describe the degree of
	      correlation between hops. Simulations with higher values of $H$ display
	      behavior closer to the Brownian limit.

		\end{quote}

		Figure 6, original caption:

	    \begin{quote}

		  The two mode model parameterizes solute behavior in the pore and tails
	      separately. We consider solutes to be within the pore region if they
	      are 0.75 nm from a given pore center, otherwise they are in the
	      tails. (a) Generally, movement is much more restricted in the tail
	      region, parameterized by lower $\sigma$ values (smaller hops) for the
	      Gaussian and L\'evy stable distributions. Values of $\alpha_h$ are
	      significantly lower for urea and acetic acid meaning there is a
	      larger probability that they will take large hops. (b) Dwell times
	      are longer in the tails. Lower values of $\alpha$ correspond to power
	      laws with heavier tails and thus higher probabilities of long dwell
	      times. There is no easily discernible trend in $\lambda$ of the
	      truncated power law distribution. Note that we used the same Hurst
	      parameter for both modes (shown in Figure~\ref{fig:hurst_barchart})
	      due to a low number of sufficiently long sequences of hops in each
	      mode. (c) Solutes spend various amounts of time in the tail and pore
	      region dependent on their size, shape and chemical functionality.
	      Methanol's small size favors occupation of the much larger accessible
	      volume in the tails. Urea and acetic acid are fairly stable in both
	      regions since they are small and hydrophilic. Ethylene glycol has a
	      slight preference for the pores likely because it is a larger
	      molecule with two hydrophilic hydroxyl groups.

        \end{quote}

	    Figure 6, modified caption:

	    \begin{quote}
		  % Figure 6 caption
		  The two mode model parameterizes solute behavior in the pore and tails
	      separately. We consider solutes to be within the pore region if they
	      are 0.75 nm from a given pore center, otherwise they are in the
	      tails. 
	      %BJC: added next two sentences
	      For (a) and (b), each bar represents the value of a single parameter, 
	      highlighted in bold, of the associated hop or dwell time distribution.
	      Parameterizations of solute motion in the pores are visualized as bars
	      in the upper half of the plots, while the parameters describing motion 
	      in the tails are reflected into the lower half of the plots.
	      %BJC: modified this sentence to cut down on detail
	      (a) Generally, movement is more restricted in the tail region, but 
	      with occasionally large hops, as implied by lower $\sigma$ (smaller
	      hops) and $\alpha_h$ (heavier tails of the hop distribution) values.
	      %for the Gaussian and L\'evy stable distributions. 
%	      Values of $\alpha_h$ are significantly lower for urea and acetic acid
%	      meaning there is a larger probability that they will take large hops.
	      (b) Dwell times are longer in the tails as implied by lower values 
	      of $\alpha$.
	      %correspond to power laws with heavier tails and thus higher probabilities of long dwell
	      %times. 
	      There is no easily discernible trend in $\lambda$ of the
	      truncated power law distribution. Note that we used the same Hurst
	      parameter for both modes (shown in Figure~\ref{fig:hurst_barchart})
	      due to a low number of sufficiently long sequences of hops in each
	      mode. (c) Solutes spend various amounts of time in the tail and pore
	      region dependent on their size, shape and chemical functionality.
	      %BJC: removing details that is already in the text
%	      Methanol's small size favors occupation of the much larger accessible
%	      volume in the tails. Urea and acetic acid are fairly stable in both
%	      regions since they are small and hydrophilic. Ethylene glycol has a
%	      slight preference for the pores likely because it is a larger
%	      molecule with two hydrophilic hydroxyl groups.

        \end{quote}

        Figure 13, modified caption:

        \begin{quote}

          The selectivity between pairs of species changes monotonically with
	      pore length. The strength of dependence on pore length depends on
	      the difference between $\beta$ values. 
		  %BJC: All of this is in the text.	      
	      \sout{The largest differences in solute
	      flux result in high selectivities at any pore length. This membrane
	      may be a good candidate for the separation of ethylene glycol from
	      acetic acid. Ethylene glycol has the lowest $\beta$ value while
	      acetic acid has the second highest, leading to strong length
	      dependence. Ethylene glycol also has the highest flux and acetic acid
	      has the lowest resulting in relatively high selectivities independent
	      of pore length.}

        \end{quote}

### 2020-06-10 Ben_Manuscripts/hdphmm/cover_letter.tex
Improvements to figure readability

	We are submitting ``\ManuscriptTitle'' by Benjamin J. Coscia, Christopher P.
	Calderon and Michael R. Shirts, for	consideration for publication as a Research
	Article in the ``Machine Learning in Physical Chemistry" special issue of the Journal 
	of Physical Chemistry, specifically in section X of % BJC: the link with section titles is an unfinished webpage
	part B of the journal. We expect this work to be of significant interest to 
	researchers who study complex dynamics in both structured and amorphous soft 
	materials. %BJC: this could be made more general. Maybe better focused for JPC though

@@

	Our work represents a considerable advancement over our most recent study 
	characterizing the same set of solute center-of-mass time series trajectories
	(Coscia and Shirts, arXiv:2004.07905 [cond-mat.soft]). % BJC: is this the right way to cite arXiv?
	In that work, we used our qualitative understanding of solute transport 
	(Coscia and Shirts \textit{J. Phys. Chem. B}, 123, 6314--6330 (2019)) in order
	to formulate stochastic models which could be used to project solute behavior
	on macroscopic time scales, thus allowing prediction of experimentally relevant
	properties like membrane selectivity. Using the HDP-AR-HMM, we can analyze solute
	trajectories with no prior knowledge of their dynamic behavior and easily produce
	models which quantitatively describe their long time scale dynamics. Since the 
	approach is non-parametric, the complexity of our data drives the complexity of
	the model. We cluster the parameter sets in order to reduce the total space of 
	dynamical states to a small set of distinct parameters which allow us to clearly
	understand physical	mechanisms in terms of solute-membrane interactions. To our 
	knowledge, this is the first application of the HDP-AR-HMM to MD data and it is
	the first time this framework has been used to gain both mechanistic and kinetic
	insight.

	%BJC: included phone numbers because it asked to do so in author's guide.
	%BJC: need at least 5 reviewers
	\noindent Some suggestions for reviewers are:
	\begin{enumerate}

	    %BJC: these are reviewers suggested for the first stochastic transport paper
	    %BJC: not sure if bad to use twice
		\item Francisco Hung has used molecular simulations to study nematic liquid crystals
		and transport under confinement within various nanostructured materials 
		(Northeastern University, 617-373-8619,\\ \href{mailto:f.hung@northeastern.edu}{f.hung@northeastern.edu}).

@@

@@

		%BJC: from structure paper
		\item Eric Jankowski has experience modeling soft materials using molecular dynamics
		including work done simulating structure factors from simulations, much like
		the techniques we use in our work (Boise State University, 208-426-5681,
		\href{mailto:ericjankowski@boisestate.edu}{ericjankowski@boisestate.edu})

		\item Emily Fox?  % she made the HDP-AR-HMM

	\end{enumerate}

>>>>

	We are submitting ``\ManuscriptTitle'' by Benjamin J. Coscia, Christopher P.
	Calderon and Michael R. Shirts, for	consideration for publication as a Research
	Article in the ``Machine Learning in Physical Chemistry" special issue of the Journal 
	of Physical Chemistry, specifically in part B, section B4: Fluid 
	Interfaces, Colloids, Polymers, Soft Matter, Surfactants, and Glassy Materials.
	We expect this work to be of significant interest to researchers who study complex 
	dynamics in both structured and amorphous soft materials.

@@

	Our work represents a considerable advancement over our most recent study 
	characterizing the same set of solute center-of-mass time series trajectories
	(Coscia and Shirts, arXiv:2004.07905 [cond-mat.soft]). In that work, we used
	our previously gained qualitative understanding of solute transport (Coscia 
	and Shirts \textit{J. Phys. Chem. B}, 123, 6314--6330 (2019)) in order to 
	formulate stochastic models which could be used to project solute behavior
	on macroscopic time scales, thus allowing prediction of experimentally relevant
	properties like membrane selectivity. Using the HDP-AR-HMM, we can analyze solute
	trajectories with no prior knowledge of their dynamic behavior and easily produce
	models which quantitatively describe their long time scale dynamics. Since the 
	approach is non-parametric, the complexity of our data drives the complexity of
	the model. We cluster the parameter sets in order to reduce the total space of 
	dynamical states to a small set of distinct parameters which allow us to clearly
	understand physical	mechanisms in terms of solute-membrane interactions. To our 
	knowledge, this is the first application of the HDP-AR-HMM to MD data and it is
	the first time this framework has been used to gain both mechanistic and kinetic
	insight.

	\noindent Some suggestions for reviewers are:
	\begin{enumerate}

		\item Francisco Hung has used molecular simulations to study nematic liquid crystals
		and transport under confinement within various nanostructured materials 
		(Northeastern University, 617-373-8619,\\ \href{mailto:f.hung@northeastern.edu}{f.hung@northeastern.edu}).

@@

		%BJC: from structure paper
		\item Eric Jankowski has experience modeling soft materials using molecular dynamics.
		(Boise State University, 208-426-5681, \href{mailto:ericjankowski@boisestate.edu}{ericjankowski@boisestate.edu})

@@

%		\item Martin Lindén has use variational Bayesian treatments of hidden Markov models in
%		order to learn from thousands of single-particle trajectories.
%		
%		\item Steve Presse
%		
	\end{enumerate}
